% ============================================
%  Article Class (This is a LaTeX2e document)
% ============================================
\documentclass[12pt]{scrartcl}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{enumitem}
\usepackage[round]{natbib}
\usepackage{color}

\newcommand\reft[3][]{#2~\ref{#3}#1}
\newcommand\refp[3][]{(#2~\ref{#3}#1)}
\newcommand\refsect[1]{\reft{Section}{#1}}
\newcommand\refsecp[1]{\refp{Sec.}{#1}}
\newcommand\reftabt[1]{\reft{Table}{#1}}
\newcommand\reftabp[1]{\refp{Tab.}{#1}}

% ======
%  Math
% ======
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathtools} % \mathclap
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{property}[thm]{Property}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{assum}[thm]{Assumption}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remarque}
\numberwithin{equation}{section}

\usepackage{amssymb}
\newcommand{\prob}[1]{\mathbb{P}\left(#1\right)}
\newcommand{\Ker}[1]{\mathrm{Ker}\left(#1\right)}
\newcommand{\Image}[1]{\mathrm{Im}\left(#1\right)}
\newcommand{\diag}[1]{\mathrm{diag}\left(#1\right)}
\newcommand{\Vect}[1]{\mathrm{Vect}\left\{#1\right\}}

% ============================
%  Figures and relative paths
% ============================
\usepackage{graphicx}
\graphicspath{{figures/}}
\usepackage{import}
\makeatletter
  \def\relativepath{\import@path}
\makeatother
\newcommand\reffigt[2][]{\reft[#1]{Figure}{#2}}
\newcommand\reffigp[2][]{\refp[#1]{Fig.}{#2}}

% ==========
%  Document
% ==========
\begin{document}

\title{rba.core\\ Building constraint matrices from RBA data.}%
\author{Biosys - MAIAGE}%
\date{\today}%

\maketitle

\newpage

\tableofcontents

\newpage

\section{Overview of RBA constraints}

In order to build an RBA system, we need to build three sets of constraints.

\begin{itemize}
  \item[($C_1$)] Mass conservation.
  \item[($C_2$)] Capacity constraints for enzymes and process machines, e.g., ribosomes.
  \item[($C_3$)] Density constraints.
\end{itemize}

All these constraints are linear (in)equalities.
We will write them in matrix form, as it is the best way to summarize
all information needed to build a constraint.

\subsection{Important concepts and conventions}

\paragraph{Production/degradation vector for macromolecules}
In the following, each macromolecule is described by its \textbf{production vector}.
It is a column vector containing the metabolites necessary to build it,
with a \textbf{minus} sign for metabolites consumed and a \textbf{plus} sign for byproducts generated.
Similarly, we define \textbf{degradation vectors}.

\paragraph{Concentration to flux conversion}
Suppose you have a metabolite at concentration $C = n/V$.
At growth-rate $\mu$, we have $dV \simeq \mu V$.
The variation of concentration due to dilution is
\[
  dC = d(n/V) = -n/V^2 dV = -\mu C
\]
In other words, in order to keep the concentration constant,
we need an input metabolite flux of
\[
  \nu = \mu C
\]
When a variable is expressed as a concentration, it may be necessary to
convert it to a flux by using this relationship.
When necessary, we indicate this conversion by inserting a conversion
matrix between the main matrix and the variable vector.

\paragraph{Variable representation}
When we go to full matrix representation in the figures,
variables are represented as a single row vector below matrices.
This convention clarifies how columns and variables are associated.
Note the transpose at the end of the row vector if you are worried about
mathematical purity.

\subsection{Variables}

RBA uses 4 sets of variables:

\begin{itemize}
  \item $\nu = (\nu_1, \ldots ,\nu_R)$: Fluxes through metabolic reactions,
  where $R$ is the number of reactions.
  \item $E = (e_1, \ldots, e_E)$: Enzyme concentrations,
  where $E$ is the number of enzymes.
  \item $P = (p_1, \ldots, p_P)$: Process machine concentrations,
  where $P$ is the number of processes.
  \item $T = (t_1, \ldots, t_T) = (TF, TC)$:
  Target values, where $T$ is the number of targets.
  Targets may be expressed as fluxes or concentrations.
  We note target fluxes $TF$ and target concentrations $TC$.
  When necessary, we assume that flux targets are all listed first,
  i.e., $T = (TF, TC)$.

\end{itemize}
Targets are production/degradation requirements for the cell to be fully functional
(e.g. keeping key metabolites at some given concentration,
producing housekeeping proteins, producing/degrading mRNAs).
Note that target values may actually be predetermined constant values,
while all others are true variables that must be optimized.
We will see later how predetermined constant values may be eliminated.

\subsection{Mathematical formulation of RBA constraints}

Before we group all constraints and variables into a single matrix,
here is an overview of the 3 family of constraints used in RBA models.

\paragraph{$(C_1)$ Mass conservation}

\[
\overbrace
{
\underbracket{S\nu}_{\parbox{2cm}{\scriptsize metabolic flux generated by metabolism}}
+
\underbracket{\mu C_E E + \mu C_P P}_{\parbox{3cm}{\scriptsize precursors used/byproducts generated by producing new molecules}}
}^{\text{\normalsize `Variable' terms}}
+
\overbrace
{
\underbracket{\mu C_{TC} TC + C_{TF}TF}_{\parbox{3cm}{\scriptsize precursors used/byproducts generated by producing new molecules}}
}^{\text{\normalsize Constant terms}} \\
=0
\]

where $S$ is the stoichiometry matrix of the metabolism,
with metabolites as rows and reactions as columns.
$C_.$ are composition matrices where every column is the composition of
one of the molecules, e.g. $(C_E)_i$ is the composition of the $i$th enzyme
(see definition of composition vector above).
Note the usage of $\mu$ along with $E$, $P$ and $TC$ as way of converting
concentrations into fluxes.

\paragraph{$(C_2)$ Capacity constraints}

\[
\diag{k_E^{backward}} E \leq \nu \leq \diag{k_E^{forward}} E
\]

where $k_E^{backward}$ is a vector containing the backward catalytic constants of enzymes.

\[
\mu MC_E E + \mu MC_P P + \mu MC_{TC} TC + MC_{TF} TF \leq \diag{k_P} P
\]

where $k_P$ is a vector containing the capacities of process machines.
$MC_.$ are matrices containing machine costs to produce macromolecules,
e.g. $(MC_E)_{ij}$ quantifies how much the machine from the $i$th process is needed
to build the $j$th enzyme.
For example, the ribosome (machine of translation process) has a processing
capacity of $k_P \simeq 72,000$ amino acid per hour.
If an enzyme contains 3 proteins of $300$ amino acid each, its machine cost
is $MC = 900$.

\paragraph{$(C_3)$ Density constraints}

\[
W_E E + W_P P + W_{TC} TC \leq \overline{D}
\]
where $W_.$ are weight matrices,
e.g. $(W_E)_{ij}$ is the weight of the $j$th enzyme in the $i$th compartement.
$\overline{D}$ is a vector containing the maximal weight per compartement,
e.g. $D_i$ is the maximum weight in the $i$th compartment.


\input{constraint_matrices}
\clearpage

\input{build_matrices}
\clearpage

\newpage
\appendix
\input{eliminate_targets}
%\clearpage

%\input{work_in_progress}

%\bibliographystyle{myplainnat}
%\bibliography{biblio}

\end{document}
% ----------------------------------------------------------------
